Clines from PDE

### 2015-10-25 notes/pde-clines.Rmd
matching sims sorta?

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
```

@@

Here's a function to solve this numerically:
```{r solve_pde_setup, cache=TRUE}
source("cline-fns.R")
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
solve.pde
cline.from.soln
```

@@

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
s <- 0.1
zone.age <- 100  # this is tau
tt <- seq(0,zone.age,length.out=100)
r <- 0.1
pde.soln <- solve.pde( u=tanh_cline, r=r, times=tt, s=s, grid=thegrid )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends="solve_pde_setup"}
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
zone.age	= 20
ndemes 		= 500
deme.size 	= 5
ninds 		= ndemes*deme.size
SIGMA 		= 10
s           = 0.1
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, depends="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+rvals)
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

OK, now let's match the PDE solution to the simulation.
First, does the cline at the selected locus match the the tanh cline?
```{r match_cline}
plot(xx,freqs[,1],type='l',xlab="spatial location",ylab="proportion")
lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red')
```
Looks pretty reasonable.

Here's the PDE solution:
```{r match_pde}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=function(...)1-tanh_cline(...), r=r, times=tt, s=s, grid=thegrid ) })
pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),])
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red',lwd=3)
    lines(1-thegrid$x.mid*SIGMA,match.clines[,k],col='green',lwd=3)
}
```

>>>>

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
source("tran.1D.R") # fix for log.A
```

@@

Here's a function to solve this numerically:
```{r solve_pde_setup, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
log_u <- function(x,...){ tanh_cline(x=-x,...,log=TRUE) }
log_1mu <- function(x,...){ tanh_cline(x=x,...,log=TRUE) }
cline.from.soln
solve.pde
```

@@

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
r <- 0.1
s <- 0.1
sigma <- 10
zone.age <- 100  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```
**Note:** the range should be wide enough these are *going to zero*
at the ends, or else we have to worry about the boundary conditions.

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
ndemes 		= 500
deme.size 	= 5
ninds 		= ndemes*deme.size
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s, pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=sigma,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, depends="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+rvals)
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

OK, now let's match the PDE solution to the simulation.
First, does the cline at the selected locus match the the tanh cline?
```{r match_cline}
plot(xx,freqs[,1],type='l',xlab="spatial location",ylab="proportion")
lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red')
```
Looks pretty reasonable.

Here's the PDE solution:
```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
}
```
